Adopting agile software development in the large organizations: A systematic literature review

### 2014-05-29 journal/article.tex
Spell check

The fact that the study is based on a quantitative basis
--> Experie reports have a subjective bias
--> The data synthesis must not rely on quantitative measures. This means that
    we also considered topics raised by only a marginal number of reports.
--> It may still be interesting to make some quantitative observations such as
    a particular problem being reported in the majority of reports.

@@

\begin{table}[h]
    \centering
    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.52\textwidth}
                     >{\raggedright\arraybackslash}p{0.40\textwidth} }
        How transformation is described   &  Primary sources  \\
        \midrule
        Narrative description of transformation process &
                P3, P5, P6, P9, P11, P12, P16, P18,
                P24, P27, P30, P36, P40, P41, P46   \\
        Application of agile methods seen from a specific viewpoint (e.g. coaching team) &
                P10, P13, P15, P19, P23, P32, P34,
                P39, P47, P48, P50, P52    \\
        Viewpoints of challenges and success factors  &
                P17, P20, P21, P29, P37, P38, P49   \\
        Discussion on establising large scale agile practices  &
                P7, P8, P31, P42, P43, P44, P45   \\
        Only brief discussion on transformation process  &
                P1, P2, P4, P14, P22, P25, P26, P28, P33, P35, P51  \\
        \bottomrule
    \end{tabular}
    \caption{The theme of each primary source, in relation to providing
             insights of the transformation process}
    \label{table:transformation_descriptiontypes}
\end{table}

@@

The pressence of a high-up manager was described to underline the organization's
commitment for change when dealing with employees or other stakeholders [P24].
The firm commitment of the high-up leader was key in keeping the agile
transformation on course in cases where objections or other problems arose [P5,
P14, P36, P46]. Another important role of the change leader was to persuade and
work for getting other people raise an interest in agile methods [P3, P31, P38,
P46, P49].

@@

Many organizations formed an agile task force or similar entity for coordinating
and leading the change. It was typical that in the beginning of the
transformation, or when it had been acknowledged that the old model had
problems, a work group was founded for coordinating the changes to be done. The
change team usually consisted of management personnel [P43, P50], and was
reinforced with external agile consultants [P18, P21, P25, P30]. Some cases
highlighted that the team had a cross organizational staff who were familiar
with the varying functions that would be affected by changes [P4, P16, P21, P41,
P48]. It was important that theere was a group of people who were strongly
committed to making a change happen [P43].

@@

Leadership for change emerged also on other organizational levels. In many cases
these change leaders were still in management positions [P14, P18, P17]. There
were also agile champions in teams who drove the change on a grassroots level
[P6, P9, P27].
As with higher-ups, these people worked for making agile methods familiar [P2,
P6, P9, P17], making change stick [P2, P18], and persuaded others to use agile
methods [P2, P9, P14].
Benefield [P6] describes that people in teams were well able to adjust the
application of the agile process as they were familiar with the problems at
hands. It was howerver hard to find people who both understood the agile
principles well, and knew how to coach within the team [P6].
% Also coaches in [P2], and a team in [P16]

@@

The agile adoption erquired coaches [P47].

@@

[P16]
* Why: To ensure trainsformation is successful
* Coach: The cross-functional rollout team

@@

It was typical to use web-based tools for project manangement [P10, P11, P16].
The use of web based tools such as wiki pages was also attributed to lessen the
effort in documentation [P14]. Other tools that were taken into use while
implementing agile methods were teleconferencing tools [P11], discussion forums
for contacting coaches [P41], and information radiators to raise the awareness
of the new method [P22, P41]. Also development and delivery infrastructure
required investments to match the needs of agile development [P50].

@@

When a general problem of frequent broken builds existed it was easy to acuire a
buy in for creating an organization-wide standard for automating builds and
tests [P11]. On the other hand, many cases emphasized that building the
continuous integration environment required a significant investment [P17, P34,
P42]. In order to have the focus and resources for creating a proper integration
environment a dedicated team was Some organizations appointed a dedicated team
to ensure the focus and resources for creating a proper integration environment
were available [P5, P16, P17].

@@

A typical practice was creating some internal material promoting the new agile
model. The material served both a documenting role and promoted awareness for
the new development model. The material was created to help explaining the agile
model to stakeholders, and to serve as a reference for teams implemeting it.
Typical in-house agile material was a manual which presented the organization
specific implementations of agile practices [P1]. Many cases described using a
wiki-based website to maintain and distribute information on agile practices
[P14, P16, P45]. Hanly et al. [P23] describes how a n organization specific
cookbook for agile implementation was created. The cookbook resolved
misconceptions, presented differences between the old model and agile, and
described agile development in traditional terms [P23]. Other agile material
organizations produced were short videos [P32], and a branded theme for the new
process [P8]. One primary source provided the organization specific Agile Policy
as an appendix [P26].

@@

Training was seen as a strong factor helping the organization to transform.
Several studies stated that training improved the chance of succeeding in the
transformation [P3, 16, P30]. It was even highlighted that the change would have
failed without transformation [P14]. As Benefield [P6] claimed: ``we saw again
and again how training could make the difference between success and failure for
a team''. Training also helped people to become more positively inclined towards
the new model [P16, P37], and at best training made people enhusiastic to change
[P3, P46]. A few cases reported that there was some disinclination to change,
and training would have helped to overcome [P11, P14].

Agile methods avoid prescribing exact ways of working but rather emphasize a
mindset of adapting to the situation. It is difficult to explain by theory how
such a mindset should be applied, and the agile practices are best learnt by
doing. Coaching teams while they apply the agile methods in practice was seen as
an important factor in change [P3, P16, P17, P30, P47]. A few cases stated even
that coaching was essential for success in transformation [P8, P18]. Also,
without coaching teams can do damage to the agile transformation if techniques
are applied improperly [P9].

@@

The choice of agile practices did affect the success of the transformation, and
also how the practices were choosen affected.

@@

Mapping the agile model to match the old model was seen as necessary in a few
cases. Although a general mapping to the old model may not be good [P44, P46],
some cases needed to preserve high level management practices [P14, P29, P48].
By allowing management to remain unchanged it was possible to introduced agile
stepwise on team level, and helped in getting management buy-in for the process
[P14, P29, P48].
In one case Scrum was considered as a wrapper for existing practices, and could
therefore be easily fit in the organization [P44]. The agile methodology was
considered complementing the existing culture, which eased management buy-in for
the new method [P44].
% Similarities and differences of agile were documented, and Scrum terminology was
% explained in terms of the old model. Mapping of management information was done
% so that for instance the information earlier found in Engineering Gantt charts
% was found in the Sprint Backlog. [P29]

@@

In a number of studies it was described that during transformation the
principles of agile should be emphasized over practices and simple mechanics
[P46, P48]. When people understand the agile values they will also understand
why the chage is being done and feel motivated [P1, P16, P42].
The people centric nature of agile development should be communicated correctly.
McDowell and Dourambeis [P32] describe how the organization launched a series of
communications events especially designed to emphasize the reasons behind the
agile practices.

@@

Cases reported that customization was part of a successful agile implementation.
For instance, teams were let to customize their practices individually, to fit
the needs of each team [P41].
Spayd [P49] writes that teams who modified agile practices to suite te large and
distributed environment ended up doing better than teams that did not.
% Some intermediate point?
To allow teams to become innovative and perform well the agile model should be
customized in a pragmatical way instead of following a strict textbook
interpretation [P10]

The customization of the agile model was also seen as an evolutionary process.
Cowan [11] suggested that in a big-bang transformation it might be necessary to
selectively compromise some parts of the agile implementation so that the core
practices can be set in place.
The agile transformation is a constantly ongoing process, and it is recommended
to regularly challenge teams to refine the agile implementation to reflect the
courrent needs of the organization [P41, P43].

@@

% ??
Especially inexperienced coaches and Scrum masters made mistakes in focusing too
much on the implementation of practces [P23, P30].

@@

Some cases had reported on challenges in requirements management, especially
highlighting difficulty in managing the gap between high level requirements and
stories handled by teams. It was stated that it is important that the
requirements are consise and small enough for the teams to handle [P5, P11].
It was recommended to invest in teaching skills in breaking down and writing
user stories [P11, P33, P48]. Gat [P17] describes how the gap between
development and requirements management was bridged by having a dedicated
Requirements Architect to help with requirements breakdown.

@@

To make a large organization change the direction of the change needs to be
communicated clearly. Atriculating urgency at start will help getting the change
under way [P49]. Feedback of changes should certainly be listened to, but in the
end it must be made clear that the old model can not be returned to [P22].
For a case where the organization was changed all at once it was reported that
the magnitude of change helped created the impression that the change was non
negotiable [P11]. Firstly, as it vas evident that a large change was to happen
people felt encouragement and permission to move on from the old way, and
secondly, the urgency eliminated wasteful discussions on wheter Scrum was a good
model or not [P11].
When the organization culture is ingrown he change vision must constantly be
reminded of, and each step towards the new model consider a victory [P48].

@@

It is clear that any positive results will facilitate an organizational
transformation. Although a recipie to fabricate positive experiences can not be
given, many cases described how they had succeeded. Benefield [P6] describes how
a survey was done to objectively measure satisfaction in agile, and the positive
results helped teams make the decision to change their way of working. Benefield
describes further how management was requesting proof on better performance as a
response to requests to increase coaching and training budget. After conducting
inquiries the coaching team was able to present an estimate of a 30\% increase
in performance, which made management highly convinced [P6]. Another case
describes how the corporate agile awareness and teaching event was designed to
be fun and engaging, which made people take more ethusiasm in applying the agile
practices [P32]. The move towards agile is assisted by making any benefits
publicly visible and celebrating even the small victories [P40, P42].

@@

Intensive communication was emphasized in a number of studies. It is important
to reach a maximum number of people throughout the organization, as without
communiction the a new model will not take root [P16, P43].
It was recommended that use of the new model is made highly visible on many
communication channels and even over-communicated [P16, P33, P41]. Mencke [P33]
summarizes the viewpoint on over-communication: ``Even if you think that your
teams understand a new method or process, repeat your communication to be
sure.'' Workshops, coaching sessions, online discussions, and newletters are
examples on different communication formats used [P41]. Another approach in
communicating the change was that managers explained and encouraged change in an
extensive series of one-to-one discussions [P22].

Enabling transparency during the transformation was another communication
related theme that was brought up, and even highlighted as a critical factor for
success [P11, P16]. Fry and Greene [P16] underline the importance of
transparency and sharing of information: ``This bias to sharing information with
everyone was critical in our ability to adapt on a daily basis to ensure our
success.'' Transparency was achieved by showing both successes and challenges
[P6], actively reaching out for feedback [P6], using project management tools to
display project status publicly [P11], by rearranging physical spaces [P42], and
by holding the meetings of the rollout team in public [P16].
By sharing experiences and status of the transformation the organization was
aligned and everyone was moving in the same direction [P42, P43].

@@

Organizing a large number of people requires some management and leadership.
Similarly, transforming the way of working of a large group requires coordinaion
and leadership [P42, P31]. In addition to the leadership provided by coaches,
specific change leaders were also mentioned. Cases indicated the importance of
having spokespersons for the change [P3, P4, P31]. Cowan [P11] describes how one
person was strongly driving the transformation, and made an indisputable
contribution for transforming the organization. Also Maples [P31] describes how
the change and agile adoption was guided by one ``counselor'' [P31]. Other cases
described that the change was lead by a competent rollout team, which had
representatives from all parts of the organization [P16, P33]. Finally, the
responsibility of line and project managers to act as change leaders was
highlighted [P18].

@@

The formation and influence of agile communities was reported to have a
significant impact on transformation.
The emergence of an agile community was reported as a success factor in the
transformation [P3]. A community was also seen as indispansable as it has power
to overcome impediments that would be too large for individuals to affect [P47].
The agile communities achieved to raisie awareness of agile methods in the
organization [P42, P41, P47], and spawned eager followers who spread the word
even further [P12].

@@

A factor in achieving change was creating alignment towards the common goal of
introducing a new development method. It was seen as an important factor that
all levels in the organization speak the same language and accept the change
[P18, P31]. Alignment was built by promoting success stories and gathering lists
of problems to tackle [P42, P46]. In one case complete alignment between teams
and within management was considered as necessary to use the agile model in a
large context [P43]. Another case highlighted creating and applying a structured
rollout process to uniformly introduce a large nuber of teams to agile [P8].

@@

The most prominent and clear success factor was having support from management.
Numerous cases made it clear that management support is an absolute necessity
[P8, P18, P27, P31, P33, P36, P43, P47, P49].
This was reflected in statemets such as ``Adopting agile, or implementing any
significant change, requires an executive’s sincere support.'' [P44],
``Executive commitment was crucial to implementing massive change.'' [P16], and
``Having upper management engaged, supportive and visible is critical for
wholesale organization involvement with Scrum.'' [P11].

@@

Visible involvment of management was also reported to motivate and encourage
employees to adopt the new model [P40]. For instance, the CTO organized training
sessions personnally [P24], and the engineering VP frequently visited sprint
demos [P11]. When the corporate level support for the agile initiative was
showing teams adopted agile methods even spontaneously [P9].

@@

In order to gain support for agile several cases highlighted the need to educate
managment [P8, P47].
Without education managers might make harmful interventions in the process, or
just stick to the old model [P16, P35, P46].
Managers not understanding the principles of agile were feeling left out with
the introduction of self organizing teams, which some times resulted in
backlashes [P6]. Providing proper training was reported to correct the
situation, and even create strong agile supporters in management [P6, P9].
Training also cleared misconceptions that management had [P29], and helped
create a consistent implementation of the agile model across the organization
[P11].

@@

Roche and Vasquez-McCall [P41] describe how prescribing the agile methods to use
lead only so-far in the transformation. The effectiveness of teaching and
communicating the transformation started to decline. To enable the
transformation to still proceed, the an organization-wide challenge was created
where teams would independently develop and showcase the best agile practices.
As a result the ownership of the methods was transfered to teams, and the new
model gained a secure foothold. [P41]

@@

An interesting success factor was the absence of a top-down mandate. Atlas [P3]
describes that the use of Scrum was spreading because teams were both allowed to
use and enabled to learn the method. The transformation itself was agile when
people felt empowerment in making the decision to adopt. The incentive to change
was created when teams learned about agile methods, and perceived that a change
would be beneficial [P3].
The absence of mandate granted genuine support for the model on the grass roots
level, which was a necessity for the process to work [P6].
It was also thought that proceeding with a top-down mandate would have made
teams use a uniform process [P41]. This would have been suboptimal, as it was
important that the practices were tailored for each team and project [P41].

@@

A few cases discussed the benefit of having change leaders without baggage of
the past. Cowan [P11] describes how the newly hired director was able to
circumvent territorial battles that existed from before, and therefore focus
fully on getting the agile model implemented. In another case external coaches
were described to better spot places for correctino in the agile model, and
their advice was received better because they were considered impartial when
being external to the organization [P16].
% Old employees were also willing to try something new, because the organization
% had gone through a time of uncertainty and restructuring [P11].

@@

Grounded theory may not be able to provide a single generally applicable model
for transformation, at least not in a literature review made based on the
current agile transformation publications. This is because of two things.
First, the organizations seem to differ too much from each other. Secondly, the
published studies focus too much on isolated phenomena within the larger
transformation.
It would require too big of an interpretation to combine all the various cases
into one theoretical model. We have chosen to present transformatoins using an
elementary two-axial classification of transformations, and listings of
individual phenomena that typically occur in a transformation.

>>>>

The fact that the study is based on a quantitative basis
--> Experience reports have a subjective bias
--> The data synthesis must not rely on quantitative measures. This means that
    we also considered topics raised by only a marginal number of reports.
--> It may still be interesting to make some quantitative observations such as
    a particular problem being reported in the majority of reports.

@@

\begin{table}[h]
    \centering
    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.52\textwidth}
                     >{\raggedright\arraybackslash}p{0.40\textwidth} }
        How transformation is described   &  Primary sources  \\
        \midrule
        Narrative description of transformation process &
                P3, P5, P6, P9, P11, P12, P16, P18,
                P24, P27, P30, P36, P40, P41, P46   \\
        Application of agile methods seen from a specific viewpoint (e.g. coaching team) &
                P10, P13, P15, P19, P23, P32, P34,
                P39, P47, P48, P50, P52    \\
        Viewpoints of challenges and success factors  &
                P17, P20, P21, P29, P37, P38, P49   \\
        Discussion on establishing large scale agile practices  &
                P7, P8, P31, P42, P43, P44, P45   \\
        Only brief discussion on transformation process  &
                P1, P2, P4, P14, P22, P25, P26, P28, P33, P35, P51  \\
        \bottomrule
    \end{tabular}
    \caption{The theme of each primary source, in relation to providing
             insights of the transformation process}
    \label{table:transformation_descriptiontypes}
\end{table}

@@

The presence of a high-up manager was described to underline the organization's
commitment for change when dealing with employees or other stakeholders [P24].
The firm commitment of the high-up leader was key in keeping the agile
transformation on course in cases where objections or other problems arose [P5,
P14, P36, P46]. Another important role of the change leader was to persuade and
work for getting other people raise an interest in agile methods [P3, P31, P38,
P46, P49].

@@

Many organizations formed an agile task force or similar entity for coordinating
and leading the change. It was typical that in the beginning of the
transformation, or when it had been acknowledged that the old model had
problems, a work group was founded for coordinating the changes to be done. The
change team usually consisted of management personnel [P43, P50], and was
reinforced with external agile consultants [P18, P21, P25, P30]. Some cases
highlighted that the team had a cross organizational staff who were familiar
with the varying functions that would be affected by changes [P4, P16, P21, P41,
P48]. It was important that there was a group of people who were strongly
committed to making a change happen [P43].

@@

Leadership for change emerged also on other organizational levels. In many cases
these change leaders were still in management positions [P14, P18, P17]. There
were also agile champions in teams who drove the change on a grassroots level
[P6, P9, P27].
As with higher-ups, these people worked for making agile methods familiar [P2,
P6, P9, P17], making change stick [P2, P18], and persuaded others to use agile
methods [P2, P9, P14].
Benefield [P6] describes that people in teams were well able to adjust the
application of the agile process as they were familiar with the problems at
hands. It was however hard to find people who both understood the agile
principles well, and knew how to coach within the team [P6].
% Also coaches in [P2], and a team in [P16]

@@

The agile adoption required coaches [P47].

@@

[P16]
* Why: To ensure transformation is successful
* Coach: The cross-functional rollout team

@@

It was typical to use web-based tools for project management [P10, P11, P16].
The use of web based tools such as wiki pages was also attributed to lessen the
effort in documentation [P14]. Other tools that were taken into use while
implementing agile methods were teleconferencing tools [P11], discussion forums
for contacting coaches [P41], and information radiators to raise the awareness
of the new method [P22, P41]. Also development and delivery infrastructure
required investments to match the needs of agile development [P50].

@@

When a general problem of frequent broken builds existed it was easy to acquire a
buy in for creating an organization-wide standard for automating builds and
tests [P11]. On the other hand, many cases emphasized that building the
continuous integration environment required a significant investment [P17, P34,
P42]. In order to have the focus and resources for creating a proper integration
environment a dedicated team was Some organizations appointed a dedicated team
to ensure the focus and resources for creating a proper integration environment
were available [P5, P16, P17].

@@

A typical practice was creating some internal material promoting the new agile
model. The material served both a documenting role and promoted awareness for
the new development model. The material was created to help explaining the agile
model to stakeholders, and to serve as a reference for teams implementing it.
Typical in-house agile material was a manual which presented the organization
specific implementations of agile practices [P1]. Many cases described using a
wiki-based website to maintain and distribute information on agile practices
[P14, P16, P45]. Hanly et al. [P23] describes how a n organization specific
cookbook for agile implementation was created. The cookbook resolved
misconceptions, presented differences between the old model and agile, and
described agile development in traditional terms [P23]. Other agile material
organizations produced were short videos [P32], and a branded theme for the new
process [P8]. One primary source provided the organization specific Agile Policy
as an appendix [P26].

@@

Training was seen as a strong factor helping the organization to transform.
Several studies stated that training improved the chance of succeeding in the
transformation [P3, 16, P30]. It was even highlighted that the change would have
failed without transformation [P14]. As Benefield [P6] claimed: ``we saw again
and again how training could make the difference between success and failure for
a team''. Training also helped people to become more positively inclined towards
the new model [P16, P37], and at best training made people enthusiastic to change
[P3, P46]. A few cases reported that there was some disinclination to change,
and training would have helped to overcome [P11, P14].

Agile methods avoid prescribing exact ways of working but rather emphasize a
mindset of adapting to the situation. It is difficult to explain by theory how
such a mindset should be applied, and the agile practices are best learned by
doing. Coaching teams while they apply the agile methods in practice was seen as
an important factor in change [P3, P16, P17, P30, P47]. A few cases stated even
that coaching was essential for success in transformation [P8, P18]. Also,
without coaching teams can do damage to the agile transformation if techniques
are applied improperly [P9].

@@

The choice of agile practices did affect the success of the transformation, and
also how the practices were chosen affected.

@@

Mapping the agile model to match the old model was seen as necessary in a few
cases. Although a general mapping to the old model may not be good [P44, P46],
some cases needed to preserve high level management practices [P14, P29, P48].
By allowing management to remain unchanged it was possible to introduced agile
step-wise on team level, and helped in getting management buy-in for the process
[P14, P29, P48].
In one case Scrum was considered as a wrapper for existing practices, and could
therefore be easily fit in the organization [P44]. The agile methodology was
considered complementing the existing culture, which eased management buy-in for
the new method [P44].
% Similarities and differences of agile were documented, and Scrum terminology was
% explained in terms of the old model. Mapping of management information was done
% so that for instance the information earlier found in Engineering Gantt charts
% was found in the Sprint Backlog. [P29]

@@

In a number of studies it was described that during transformation the
principles of agile should be emphasized over practices and simple mechanics
[P46, P48]. When people understand the agile values they will also understand
why the change is being done and feel motivated [P1, P16, P42].
The people centric nature of agile development should be communicated correctly.
McDowell and Dourambeis [P32] describe how the organization launched a series of
communications events especially designed to emphasize the reasons behind the
agile practices.

@@

Cases reported that customization was part of a successful agile implementation.
For instance, teams were let to customize their practices individually, to fit
the needs of each team [P41].
Spayd [P49] writes that teams who modified agile practices to suite the large and
distributed environment ended up doing better than teams that did not.
% Some intermediate point?
To allow teams to become innovative and perform well the agile model should be
customized in a pragmatical way instead of following a strict textbook
interpretation [P10]

The customization of the agile model was also seen as an evolutionary process.
Cowan [11] suggested that in a big-bang transformation it might be necessary to
selectively compromise some parts of the agile implementation so that the core
practices can be set in place.
The agile transformation is a constantly ongoing process, and it is recommended
to regularly challenge teams to refine the agile implementation to reflect the
current needs of the organization [P41, P43].

@@

% ??
Especially inexperienced coaches and Scrum masters made mistakes in focusing too
much on the implementation of practices [P23, P30].

@@

Some cases had reported on challenges in requirements management, especially
highlighting difficulty in managing the gap between high level requirements and
stories handled by teams. It was stated that it is important that the
requirements are concise and small enough for the teams to handle [P5, P11].
It was recommended to invest in teaching skills in breaking down and writing
user stories [P11, P33, P48]. Gat [P17] describes how the gap between
development and requirements management was bridged by having a dedicated
Requirements Architect to help with requirements breakdown.

@@

To make a large organization change the direction of the change needs to be
communicated clearly. Articulating urgency at start will help getting the change
under way [P49]. Feedback of changes should certainly be listened to, but in the
end it must be made clear that the old model can not be returned to [P22].
For a case where the organization was changed all at once it was reported that
the magnitude of change helped created the impression that the change was non
negotiable [P11]. Firstly, as it was evident that a large change was to happen
people felt encouragement and permission to move on from the old way, and
secondly, the urgency eliminated wasteful discussions on whether Scrum was a good
model or not [P11].
When the organization culture is ingrown he change vision must constantly be
reminded of, and each step towards the new model consider a victory [P48].

@@

It is clear that any positive results will facilitate an organizational
transformation. Although a recipe to fabricate positive experiences can not be
given, many cases described how they had succeeded. Benefield [P6] describes how
a survey was done to objectively measure satisfaction in agile, and the positive
results helped teams make the decision to change their way of working. Benefield
describes further how management was requesting proof on better performance as a
response to requests to increase coaching and training budget. After conducting
inquiries the coaching team was able to present an estimate of a 30\% increase
in performance, which made management highly convinced [P6]. Another case
describes how the corporate agile awareness and teaching event was designed to
be fun and engaging, which made people take more enthusiasm in applying the agile
practices [P32]. The move towards agile is assisted by making any benefits
publicly visible and celebrating even the small victories [P40, P42].

@@

Intensive communication was emphasized in a number of studies. It is important
to reach a maximum number of people throughout the organization, as without
communication the a new model will not take root [P16, P43].
It was recommended that use of the new model is made highly visible on many
communication channels and even over-communicated [P16, P33, P41]. Mencke [P33]
summarizes the viewpoint on over-communication: ``Even if you think that your
teams understand a new method or process, repeat your communication to be
sure.'' Workshops, coaching sessions, online discussions, and newsletters are
examples on different communication formats used [P41]. Another approach in
communicating the change was that managers explained and encouraged change in an
extensive series of one-to-one discussions [P22].

Enabling transparency during the transformation was another communication
related theme that was brought up, and even highlighted as a critical factor for
success [P11, P16]. Fry and Greene [P16] underline the importance of
transparency and sharing of information: ``This bias to sharing information with
everyone was critical in our ability to adapt on a daily basis to ensure our
success.'' Transparency was achieved by showing both successes and challenges
[P6], actively reaching out for feedback [P6], using project management tools to
display project status publicly [P11], by rearranging physical spaces [P42], and
by holding the meetings of the roll-out team in public [P16].
By sharing experiences and status of the transformation the organization was
aligned and everyone was moving in the same direction [P42, P43].

@@

Organizing a large number of people requires some management and leadership.
Similarly, transforming the way of working of a large group requires coordination
and leadership [P42, P31]. In addition to the leadership provided by coaches,
specific change leaders were also mentioned. Cases indicated the importance of
having spokespersons for the change [P3, P4, P31]. Cowan [P11] describes how one
person was strongly driving the transformation, and made an indisputable
contribution for transforming the organization. Also Maples [P31] describes how
the change and agile adoption was guided by one ``counselor'' [P31]. Other cases
described that the change was lead by a competent roll-out team, which had
representatives from all parts of the organization [P16, P33]. Finally, the
responsibility of line and project managers to act as change leaders was
highlighted [P18].

@@

The formation and influence of agile communities was reported to have a
significant impact on transformation.
The emergence of an agile community was reported as a success factor in the
transformation [P3]. A community was also seen as indispensable as it has power
to overcome impediments that would be too large for individuals to affect [P47].
The agile communities achieved to raise awareness of agile methods in the
organization [P42, P41, P47], and spawned eager followers who spread the word
even further [P12].

@@

A factor in achieving change was creating alignment towards the common goal of
introducing a new development method. It was seen as an important factor that
all levels in the organization speak the same language and accept the change
[P18, P31]. Alignment was built by promoting success stories and gathering lists
of problems to tackle [P42, P46]. In one case complete alignment between teams
and within management was considered as necessary to use the agile model in a
large context [P43]. Another case highlighted creating and applying a structured
roll-out process to uniformly introduce a large number of teams to agile [P8].

@@

The most prominent and clear success factor was having support from management.
Numerous cases made it clear that management support is an absolute necessity
[P8, P18, P27, P31, P33, P36, P43, P47, P49].
This was reflected in statements such as ``Adopting agile, or implementing any
significant change, requires an executive’s sincere support.'' [P44],
``Executive commitment was crucial to implementing massive change.'' [P16], and
``Having upper management engaged, supportive and visible is critical for
wholesale organization involvement with Scrum.'' [P11].

@@

Visible involvement of management was also reported to motivate and encourage
employees to adopt the new model [P40]. For instance, the CTO organized training
sessions personally [P24], and the engineering VP frequently visited sprint
demos [P11]. When the corporate level support for the agile initiative was
showing teams adopted agile methods even spontaneously [P9].

@@

In order to gain support for agile several cases highlighted the need to educate
management [P8, P47].
Without education managers might make harmful interventions in the process, or
just stick to the old model [P16, P35, P46].
Managers not understanding the principles of agile were feeling left out with
the introduction of self organizing teams, which some times resulted in
backlashes [P6]. Providing proper training was reported to correct the
situation, and even create strong agile supporters in management [P6, P9].
Training also cleared misconceptions that management had [P29], and helped
create a consistent implementation of the agile model across the organization
[P11].

@@

Roche and Vasquez-McCall [P41] describe how prescribing the agile methods to use
lead only so-far in the transformation. The effectiveness of teaching and
communicating the transformation started to decline. To enable the
transformation to still proceed, the an organization-wide challenge was created
where teams would independently develop and showcase the best agile practices.
As a result the ownership of the methods was transferred to teams, and the new
model gained a secure foothold. [P41]

@@

An interesting success factor was the absence of a top-down mandate. Atlas [P3]
describes that the use of Scrum was spreading because teams were both allowed to
use and enabled to learn the method. The transformation itself was agile when
people felt empowerment in making the decision to adopt. The incentive to change
was created when teams learned about agile methods, and perceived that a change
would be beneficial [P3].
The absence of mandate granted genuine support for the model on the grass roots
level, which was a necessity for the process to work [P6].
It was also thought that proceeding with a top-down mandate would have made
teams use a uniform process [P41]. This would have been sub-optimal, as it was
important that the practices were tailored for each team and project [P41].

@@

A few cases discussed the benefit of having change leaders without baggage of
the past. Cowan [P11] describes how the newly hired director was able to
circumvent territorial battles that existed from before, and therefore focus
fully on getting the agile model implemented. In another case external coaches
were described to better spot places for correction in the agile model, and
their advice was received better because they were considered impartial when
being external to the organization [P16].
% Old employees were also willing to try something new, because the organization
% had gone through a time of uncertainty and restructuring [P11].

@@

Grounded theory may not be able to provide a single generally applicable model
for transformation, at least not in a literature review made based on the
current agile transformation publications. This is because of two things.
First, the organizations seem to differ too much from each other. Secondly, the
published studies focus too much on isolated phenomena within the larger
transformation.
It would require too big of an interpretation to combine all the various cases
into one theoretical model. We have chosen to present transformations using an
elementary two-axial classification of transformations, and listings of
individual phenomena that typically occur in a transformation.
